Testing atlantisom: generate census length comps for California Current Atlantis

### 2019-05-11 visualization_Rmd/CCATrueLengthCompTest.Rmd
length comp vis CCA

---
title: "Testing atlantisom: generate census length comps"
author: "Sarah Gaichas and Christine Stawitz"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
bibliography: "packages.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'knitr', 'rmarkdown', 'tidyr', 'dplyr', 'ggplot2',
  'data.table', 'here', 'ggforce', 'ggthemes'
), 'packages.bib')
```

@@

This section uses the `atlantisom::create_survey()` and `atlantisom::sample_fish()` to get a biological sample dataset. From that dataset, several other functions can be run:

 *  age samples from sample_ages
 *  length samples from sample_lengths (TO BE WRITTEN)
 *  weight samples from sample_weights (TO BE WRITTEN)
 *  diet samples from sample_diet

Atlantis outputs numbers by cohort (stage-age) and growth informtation, but does not output size of animals directly. The function we are testing here, `atlantisom::calc_age2length` converts numbers by cohort to a length composition. The [workflow originally envisioned](https://onedrive.live.com/?authkey=%21AFQkOoKRz64TLUw&cid=59547B4CB95EF108&id=59547B4CB95EF108%216291&parId=59547B4CB95EF108%216262&o=OneUp) was to create a survey, sample fish from the survey, then apply this function. We determined that it will not work better to create a length comp for the whole population, which could then be sampled. We rethought what we needed [here](https://sgaichas.github.io/poseidon-dev/RethinkSamplingFunctions.html) and test the new functions in this document.

### This stuff is biolerplate from other docs, repeated here for completeness: 
To create a survey, the user specifies the timing of the survey, which species are captured, the spatial coverage of the survey, the species-specific survey efficiency ("q"), and the selectivity at age for each species.

@@

```{r stdsurveyNbased}

@@

Next, get true annual (cohort) age comp from this census survey based on run truth. (is there a standard Atlantis output I can compare this to as we did for biomass?)

@@

However. We need both numbers and weights to get the length comp, and ultimately the weight at age for use in assessments.

After using functions without thinking first and then [rethinking sampling flow](https://sgaichas.github.io/poseidon-dev/RethinkSamplingFunctions.html), we will now test new functons to aggregate both densities and numbers appropriately for input into the biological sampling workflow.

What we want is to subset the nums, resn, and structn components of run_truth using the same create_survey design selecting boxes, time, and species. We need only apply survey efficiency and selectivity to nums, however. So perhaps the new aggregateDensityData function can be applied to resn and structn?

```{r aggdens}

@@

To do the proper aggregation and not apply the multinomial sampling to the density data, I rewrote `sample_fish` to apply the median if `sample=FALSE` in the function call. Test this:

```{r censussurvey-calcage2length, warning=FALSE, message=FALSE}

@@

# now cut these down to a single species for testing
# this should still represent a census but with polygon and layer aggregated

atf_numsallhigh <- numsallhigh[numsallhigh$species == "Arrowtooth_flounder",]
atf_structnall <- structnall[structnall$species == "Arrowtooth_flounder",]
atf_resnall <- resnall[resnall$species == "Arrowtooth_flounder",]

atf_length_censussurvsamp <- calc_age2length(structn = atf_structnall,
                                 resn = atf_resnall,
                                 nums = atf_numsallhigh,
                                 biolprm = truth$biolprm, fgs = truth$fgs,
                                 CVlenage = 0.1, remove.zeroes=TRUE)

@@

Maybe this will work, we can hope. It looks like it did...

```{r atflengthsamp1-test}

lfplot <- ggplot(atf_length_censussurvsamp$natlength, aes(upper.bins)) +
  geom_bar(aes(weight = atoutput)) +
  theme_tufte()

@@

## Is this working??

OK, maybe we have something here. Next I need to see how long it takes to run for all species in the survey.

```{r allspplength, eval=FALSE}

@@

Answer: this ran for 5.5 hours. But it was really only 5 hours 20 minutes that was the last code block! Saving the file is highly recommended.

>>>>

---
title: "Testing atlantisom: generate census length comps for California Current Atlantis"
author: "Sarah Gaichas and Christine Stawitz"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
bibliography: "packages.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'knitr', 'rmarkdown', 'tidyr', 'dplyr', 'ggplot2',
  'data.table', 'here', 'ggforce', 'ggthemes'
), 'packages.bib')
```

@@

Full methods are explained [here](https://sgaichas.github.io/poseidon-dev/TrueLengthCompTest.html).

@@

```{r stdsurveyNbased, echo=TRUE}

@@

@@

The following gets both numbers and weights to get the length comp, and ultimately the weight at age for use in assessments.

First we subset the nums, resn, and structn components of `run_truth` using the same `create_survey` design selecting boxes, time, and species. We need only apply survey efficiency and selectivity to nums, however (done above), so we apply `aggregateDensityData` here.

```{r aggdens, echo=TRUE}

@@

To do the proper aggregation and not apply the multinomial sampling to the density data, I rewrote `sample_fish` to apply the median if `sample=FALSE` in the function call.

```{r censussurvey-samplefish, warning=FALSE, message=FALSE, echo=TRUE}

@@

@@

Select a single species for testing. Generating length composition for a single species should take just a few minutes to run.

```{r singlespp-lengthcomp, echo=TRUE}

# this should still represent a census but with polygon and layer aggregated
# WARNING: selecting a species by name is hardcoded for a particular model

ss_numsallhigh <- numsallhigh[numsallhigh$species == "Yelloweye_rockfish",]
ss_structnall <- structnall[structnall$species == "Yelloweye_rockfish",]
ss_resnall <- resnall[resnall$species == "Yelloweye_rockfish",]

ss_length_censussurvsamp <- calc_age2length(structn = ss_structnall,
                                 resn = ss_resnall,
                                 nums = ss_numsallhigh,
                                 biolprm = truth$biolprm, fgs = truth$fgs,
                                 CVlenage = 0.1, remove.zeroes=TRUE)

This gives annual length comps:

```{r ss-lengthsamp1-test}

lfplot <- ggplot(ss_length_censussurvsamp$natlength, aes(upper.bins)) +
  geom_bar(aes(weight = atoutput)) +
  theme_tufte()

@@

## Run for all species then SAVE

This code block can take hours (5+ for CCA) to run, so be prepared, but then the saved file can be read back in to have length comps for all surveyed species.

```{r allspplength, eval=FALSE, echo=TRUE}

@@

Read it back in:

```{r loadlengthcomp, echo=TRUE}

length_censussurvsamp <- readRDS(file.path(d.name, paste0(scenario.name, "length_censussurvsamp.rds")))

Demo plots for other species:

```{r moreplots, echo=TRUE}

ss_length_censussurvsamp <- length_censussurvsamp[length_censussurvsamp$species == "Herring",]

lfplot <- ggplot(ss_length_censussurvsamp$natlength, aes(upper.bins)) +
  geom_bar(aes(weight = atoutput)) +
  theme_tufte()

So this looks like our best approximation of a true length comp for a particular Atlantis run. Comparisons with standard surveys next.
